Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-27 drop.tex
Edits to dropping section.

Figure~\ref{dummy} shows the overheads of array bounds check with filtering
based on initialized metadata (see Section~\ref{sec:evaluation} for details on
the experimental methodology). The overheads range from XX\% to XX\%. In some
contexts, this overhead may still be considered to be too high. In this
section, we discuss how we can selectively perform monitoring in order to
further reduce overheads and match a specified overhead budget.

The high-level idea is to drop monitoring operations in order to remain within
the specified overhead budget. Given an overhead budget, the system estimates
the run-time overhead caused by monitoring (Section~\ref{sec:drop.slack}). When
this overhead exceeds the budget, then the system drops monitoring operations.
In order to prevent false positives, we must be able to identify flows that
have been dropped. This can be done by extending the dataflow engine to keep
track of an additional \emph{invalid flag} (Section~\ref{sec:drop.false_positives}).

@@

% Slack
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/slack.pdf}
    \vspace{-0.3in}
    \caption{Slack and its effect on monitoring over time.}
    \label{fig:drop.slack}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% Slack tracking module
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/stm.pdf}
    \vspace{-0.3in}
    \caption{Slack tracking and drop decision hardware.}
    \label{fig:drop.stm}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

% Instead, we develop a model to estimate the execution time overhead budget in a
% gradual manner and without the need to modify the main program. Specifically,
We specify the overhead budget as a percentage of the main program's execution
cycles without monitoring. We define \emph{slack} as the number of cycles of
monitoring overhead that can be incurred while staying within the budget
target. Slack is essentially the difference between the actual overheads seen
and the budget specified. Slack is generated as the main program runs.  For
example, if no monitoring overheads occur during 1000 cycles of the main
program's execution and the designer sets a 20\% overhead target, then the
slack that is built up during this period is 200 cycles. This slack can be used
to perform monitoring without exceeding the overhead budget.
Figure~\ref{fig:drop.slack} shows an example of how slack can change over time.
As monitoring overheads occur, the slack is decremented. 
% If the slack falls below the overhead expected for handling a monitoring event,
% then the event is dropped. 
If the slack falls below zero (i.e., the overhead budget is exceeded), then
monitoring events are dropped.
In addition to this accumulated slack, a small amount of initial slack
can be given in order for monitoring to be performed at the start of a program.

Figure~\ref{fig:drop.stm} shows a hardware slack tracking module (STM) for
keeping track of slack. The STM uses a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, slack
is decremented.

@@

% Flag propagation rules for BC
\begin{table}[tb]
  \begin{center}
    \begin{small}
    \input{tabs/propagation_rules}
    \end{small}
    \caption{Flag propagation rules for array bounds check. X indicates that no
    propagation is done while 0 or 1 indicates that the destination flag is set
    to that value.}
    \label{tab:drop.propagation_rules}
    \vspace{-0.2in}
  \end{center}
\end{table}

We extend the dataflow engine to keep track of invalidation flags in addition
to the clean flags discussed in Section~\ref{sec:filter}.
Since metadata derived from dropped metadata is also invalid. We propagate the
drop flag in a similar manner to the clean flags.  In the typical case, the
propagation rule is
that if any of the source operands is marked as invalid, then the destination
operand is marked as dropped. This is slightly different than the
policy for clean flags which propagates a clean flag only if both
sources are marked as clean. Table~\ref{tab:drop.propagation_rules} shows the
propagation rules that we use for array bounds check.  However, we note that
the propagation rules can be customized depending on the monitor by
reconfiguring the Filter Decision Table.
Similarly to filtering for clean flags, we can also filter out
useless monitoring for invalid metadata.
Thus, we see that by tracking two bits of information, we are able to use the
dataflow engine to both filter out monitoring operations for clean metadata as
well as selectively perform monitoring without false positives.

\subsection{Dropping Decision}

% Varying slack's impact on coverage for UMC
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf}
    \vspace{-0.3in}
    \caption{Example dependence graph for metadata. Blue (dark) nodes represent
    events where checks are performed.}
    \label{fig:drop.dataflow_graph}
    \vspace{-0.1in}
  \end{center}
\end{figure}

One of the issues with waiting until the overhead budget is reached (i.e.,
slack is zero) to start dropping monitoring events is that this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:drop.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be also dropped. In the
example, we see that if event {\tt E} is dropped, then this meant that
monitoring for events {\tt C} and {\tt D} was wasted. The results of those
monitoring operations were not used for any monitoring checks.

One method to eliminate this wasted work is to only make dropping decisions at
the root of these metadata flow graphs. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source dropping} and we refer to the previous policy of waiting until
the overhead budget is reached as \emph{last-minute dropping}. Source dropping will
result in no wasted work. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. Source dropping is expected
to work well when there are a large number of independent metadata flows.

Another possible method is to make the dropping decisions somewhere in between
\emph{source dropping} and \emph{last-minute dropping}. From
Figure~\ref{fig:drop.dataflow_graph}, what we would like to do is make a
decision at event {\tt C} whether to perform monitoring on that branch of the
metadata flow. If we choose to perform monitoring, then we want to perform all
monitoring on that flow. Instead, if we choose to drop event {\tt C}, then the
entire flow is dropped with no wasted work. Similarly, dropping event {\tt F}
drops the bottom flow with no wasted work and event {\tt A} can be dropped to
drop the entire shown metadata flow with no wasted work. If the program can be
analyzed or profiled to identify these instructions which, if dropped, lead to
no wasted work, then this information can be used at run-time to minimize the
amount of wasted work. We refer to this policy as \emph{min-point dropping}.
Min-point dropping is a coarser-grained decision than last-minute dropping, but
finer-grained that source dropping. Thus, its ability to match overheads is
expected to sit between these other two policies. Similarly, there is still a
chance of wasted work due to strange splitting and merging metadata flow graph
shapes but typically we expect much less wasted work than last-minute
dropping. Thus, we expect the coverage achieved by min-point dropping for a
certain overhead to fall between source and last-minute dropping. These three
dropping policies create a trade-off space between coverage achieved and
ability to match an overhead target which we evaluate in detail in
Section~\ref{sec:evaluation}.

\subsection{Multiple-Run Coverage}
Coverage can be considered for a single or average run instance. It can also be
measured over multiple users/runs in order to capture total bug coverage of a
program.

The system presented so far operates in a deterministic manner and so each run
will perform the same monitoring checks. However, we can introduce randomness
in order to achieve different dropping patterns over runs and achieve higher
multiple-run coverage.

Due to the multiplicative effect of randomness, long dependence flows with
multiple points where randomness is introduced leads to a small chance that the
entire flow is monitored. This is a significant issue for last-minute dropping
where all events could be dropped. Min-point dropping reduces this effect by
reducing the points where dropping decisions are made while source-dropping
eliminates this issue entirely.

\subsection{Multi-Core Implementation}

This can be extended to mulit-core by using multiple dataflow engines and a
similar coherency protocol as that used for data caches.

>>>>

Figure~\ref{dummy} shows the overheads of array bounds check with filtering
based on clean metadata (see Section~\ref{sec:evaluation} for details on
the experimental methodology). The overheads range from XX\% to XX\%. In some
contexts, this overhead may still be considered to be too high, forcing
monitoring to be disabled. Instead, it would be useful if partial monitoring
can be done in order to provide partial protection with further reduced
overheads. In this
section, we discuss how we can selectively perform monitoring in order to
further reduce overheads.

% Overview of dropping architecture
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/drop_overview.pdf}
    \vspace{-0.3in}
    \caption{Block diagram of architecture for monitoring with adjustable overheads.}
    \label{fig:drop.overview}
    \vspace{-0.1in}
  \end{center}
\end{figure}

The goal is to be able to perform as much monitoring as possible within a specified overhead budget.
The high-level approach is to keep track of monitoring overheads at run-time
and drop monitoring operations when the overhead exceed the specified budget.
A block diagram of the architecture we will present in this section is shown in
Figure~\ref{fig:drop.overview}.  In order to do limit the overheads of
monitoring, we need to first be able to track the gap between the specified
overhead budget and the monitoring overhead seen at run-time. We refer to this
as slack and discuss how we measure it in Section~\ref{sec:drop.slack}. 
Dropping monitoring operations can cause false positives which we prevent by
extending the dataflow engine to track and filter out dropped monitoring flows (see
Section~\ref{sec:drop.false_positives}).
Given that we know how much slack we have for monitoring and can drop
monitoring operations without false positives, we must decide when and which
monitoring operations to drop.  There exists some interesting trade-offs in the
policy for this dropping decision which we discuss in
Section~\ref{sec:drop.policies}.

@@

% Slack
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/slack.pdf}
    \vspace{-0.3in}
    \caption{Slack and its effect on monitoring over time.}
    \label{fig:drop.slack}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% Slack tracking module
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/stm.pdf}
    \vspace{-0.2in}
    \caption{Slack tracking and drop decision hardware.}
    \label{fig:drop.stm}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

% Instead, we develop a model to estimate the execution time overhead budget in a
% gradual manner and without the need to modify the main program. Specifically,
We specify the overhead budget as a percentage of the main program's execution
cycles without monitoring. We define \emph{slack} as the number of cycles of
monitoring overhead that can be incurred while staying within the budget
target. Slack is essentially the difference between the actual overheads seen
and the budget specified. Slack is generated as the main program runs and consumed as monitoring overheads occur.  For
example, if no monitoring overheads occur during 1000 cycles of the main
program's execution and the designer sets a 20\% overhead target, then the
slack that is built up during this period is 200 cycles. If the main core is
then stalled for 50 cycles due to monitoring, then the remaining slack is 150
cycles. 
In addition to this accumulated slack, a small amount of initial slack
can be given in order for monitoring to be performed at the start of a program.
Figure~\ref{fig:drop.slack} shows an example of how slack can change over time.
If the slack falls below zero (i.e., the overhead budget is exceeded), then
monitoring events should be dropped.

Figure~\ref{fig:drop.stm} shows a hardware slack tracking module for
keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle. The value of this counter is the
accumulated slack. Whenever the main core is stalled due to the monitor, slack
is decremented.

@@

% Flag propagation rules for BC
\begin{table}[tb]
  \begin{center}
    \begin{small}
    \input{tabs/propagation_rules}
    \end{small}
    \caption{Flag propagation rules for array bounds check. A dash indicates that no
    propagation is done while 0 or 1 indicates that the destination flag is set
    to that value.}
    \label{tab:drop.propagation_rules}
    \vspace{-0.2in}
  \end{center}
\end{table}

We can track these invalid flags by extending the dataflow engine presented in
Section~\ref{sec:filter} to keep track of an additional bit for each flag. That
is, we use the dataflow engine to track 2-bit flags in order to keep track of
both clean/dirty information as well as valid/invalid information.
Similar to filtering for clean flags, we can also filter out
monitoring corresponding to invalid metadata which would be useless to perform.
However, it is important that we propagate the
invalidation flag in these circumstances so that the entire flow is marked as invalid. 
In the typical case, the propagation rule for this is
that if any of the source operands is marked as invalid, then the destination
operand is marked as invalid. This is slightly different than the
policy for clean flags which propagates a clean flag only if both
sources are marked as clean. Table~\ref{tab:drop.propagation_rules} shows the
propagation rules that are used for array bounds check. However, we note that
the propagation rules can be customized depending on the monitor by
reconfiguring the filter decision table.

\subsection{Dropping Decision}
\label{sec:drop.policies}

% Varying slack's impact on coverage for UMC
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf}
    \vspace{-0.3in}
    \caption{Example dependence graph for metadata. Square nodes represent
    events where checks are performed.} 
    \label{fig:drop.dataflow_graph}
    \vspace{-0.1in}
  \end{center}
\end{figure}

The simplest policy of when to drop monitoring events is to drop all monitoring events as
long as slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:drop.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be also dropped. Square
nodes indicate events where monitoring checks are performed. In the
example, event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results not used for any monitoring checks.

\begin{figure}
  \begin{center}
  \subfloat[all-event dropping]{
    \includegraphics[width=\columnwidth]{figs/all_drop.pdf}
    \label{fig:drop.all_drop}
  }
  \vspace{-0.1in}
  \\
  \subfloat[source-event dropping]{
    \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
    \label{fig:drop.source_drop}
  }
  \vspace{-0.1in}
  \\
  \subfloat[min-event dropping]{
    \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
    \label{fig:drop.min_drop}
  }
  \end{center}
  \vspace{-0.2in}
  \caption{Comparison of dropping policies. Blue (dark) nodes indicate where
  dropping decisions are made.}
  \label{fig:drop.policies}
\end{figure}

One method to eliminate this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-event dropping} and we refer to the previous policy of waiting until
the overhead budget is reached as \emph{all-event dropping}.
Figures~\ref{fig:drop.all_drop} and \ref{fig:drop.source_drop} shows a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. Source dropping is expected
to work well when there are a large number of independent metadata flows.

Another possible method is to make the dropping decisions somewhere in between
source-event dropping and all-event dropping. Looking at the graph from 
Figure~\ref{fig:drop.dataflow_graph}, what we would like to do is make a
decision at event {\tt C} whether to perform monitoring on that branch of the
metadata flow. If we choose to perform monitoring, then we want to perform all
monitoring on that flow. Instead, if we choose to drop event {\tt C}, then the
entire flow is dropped with no wasted work. Similarly, dropping event {\tt F}
drops the bottom flow with no wasted work and event {\tt A} can be dropped to
drop the entire shown metadata flow with no wasted work. If the program can be
analyzed or profiled to identify these instructions which, if dropped, lead to
no wasted work, then this information can be used at run-time to minimize the
amount of wasted work. We refer to this policy as \emph{min-event dropping}.
Figure~\ref{fig:drop.min_drop} shows the points were dropping decisions are
made for this policy.
Min-event dropping is a coarser-grained decision than all-event dropping, but
finer-grained that source dropping. Thus, its ability to match overheads is
expected to sit between these other two policies. Similarly, there is still a
chance of wasted work due to strange splitting and merging metadata flow graph
shapes, but typically we expect much less wasted work than all-event 
dropping. Thus, we expect the coverage achieved by min-event dropping for a
certain overhead to fall between source and all-event dropping. These three
dropping policies create a trade-off space between coverage achieved and
ability to match an overhead target which we evaluate in detail in
Section~\ref{sec:evaluation}.

\textbf{[We probably need more details of how min-event dropping is
implemented. How is analysis done? How is the information conveyed to
run-time?]}

% \subsection{Multiple-Run Coverage}
% Coverage can be considered for a single or average run instance. It can also be
% measured over multiple users/runs in order to capture total bug coverage of a
% program.
% 
% The system presented so far operates in a deterministic manner and so each run
% will perform the same monitoring checks. However, we can introduce randomness
% in order to achieve different dropping patterns over runs and achieve higher
% multiple-run coverage.
% 
% Due to the multiplicative effect of randomness, long dependence flows with
% multiple points where randomness is introduced leads to a small chance that the
% entire flow is monitored. This is a significant issue for last-minute dropping
% where all events could be dropped. Min-point dropping reduces this effect by
% reducing the points where dropping decisions are made while source-dropping
% eliminates this issue entirely.
% 
% \subsection{Multi-Core Implementation}
% 
% This can be extended to mulit-core by using multiple dataflow engines and a
% similar coherency protocol as that used for data caches.
